% arara: xelatex
\documentclass[12pt]{article}

\usepackage{tikz} % картинки в tikz
\usepackage{microtype} % свешивание пунктуации

\usepackage{array} % для столбцов фиксированной ширины

\usepackage{indentfirst} % отступ в первом параграфе

\usepackage{sectsty} % для центрирования названий частей
\allsectionsfont{\centering}

\usepackage{amsmath, amssymb, amsthm} % куча стандартных математических плюшек

\usepackage{amsfonts}

\usepackage{comment}

\usepackage[top=2cm, left=1.2cm, right=1.2cm, bottom=2cm]{geometry} % размер текста на странице

\usepackage{lastpage} % чтобы узнать номер последней страницы

\usepackage{enumitem} % дополнительные плюшки для списков
%  например \begin{enumerate}[resume] позволяет продолжить нумерацию в новом списке
\usepackage{caption}


\usepackage{hyperref} % гиперссылки

\usepackage{multicol} % текст в несколько столбцов


\usepackage{fancyhdr} % весёлые колонтитулы
\pagestyle{fancy}
\lhead{ICEF, Stochastic Calculus 2022}
\chead{DON'T PANIC}
\rhead{2022-12-29}
\lfoot{120 minutes, A4 cheat sheet and calculator are ok, $W_t$ denotes a Wiener process,
10 points for each problem}
\cfoot{}
\rfoot{}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

\let\P\relax
\DeclareMathOperator{\P}{\mathbb{P}}
\DeclareMathOperator{\Cov}{\mathbb{C}ov}
\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Var}{\mathbb{V}ar}
\newcommand{\cN}{\mathcal{N}}

\usepackage{todonotes} % для вставки в документ заметок о том, что осталось сделать
% \todo{Здесь надо коэффициенты исправить}
% \missingfigure{Здесь будет Последний день Помпеи}
% \listoftodos - печатает все поставленные \todo'шки


% более красивые таблицы
\usepackage{booktabs}
% заповеди из докупентации:
% 1. Не используйте вертикальные линни
% 2. Не используйте двойные линии
% 3. Единицы измерения - в шапку таблицы
% 4. Не сокращайте .1 вместо 0.1
% 5. Повторяющееся значение повторяйте, а не говорите "то же"



\usepackage{fontspec}
\usepackage{polyglossia}

\setmainlanguage{english}
\setotherlanguages{russian}

% download "Linux Libertine" fonts:
% http://www.linuxlibertine.org/index.php?id=91&L=1
\setmainfont{Linux Libertine O} % or Helvetica, Arial, Cambria
% why do we need \newfontfamily:
% http://tex.stackexchange.com/questions/91507/
\newfontfamily{\cyrillicfonttt}{Linux Libertine O}

\AddEnumerateCounter{\asbuk}{\russian@alph}{щ} % для списков с русскими буквами
% \setlist[enumerate, 2]{label=\asbuk*),ref=\asbuk*}




\begin{document}

\begin{enumerate}

%    \item Consider $X_t = \int_0^t W_u^3 dW_u + \int_0^t (W_u^3 + 3W_u u ) du - W_t^3 \cdot t$.

%\begin{enumerate}
%    \item Find $dX_t$ and the corresponding full form. 
%    \item Is $X_t$ a martingale?
% \end{enumerate}

\item Consider $X_t = \exp(-2W_t - 2t)$.
\begin{enumerate}
    \item Find $dX_t$. Write the corresponding full form. Is $X_t$ a martingale?
    \item Find $\E(X_t)$ and $\Var(X_t)$.
    \item Find $\int_0^t X_u dW_u$.
\end{enumerate}


\item Let $(W_t)$ be a Wiener process.
\begin{enumerate}
    \item Find $\E(W_5 W_4 \mid W_4)$, $\Var(W_5 W_4 \mid W_4)$.
    \item Find $\E(W_5 W_4 W_3 \mid W_4)$, $\Var(W_5 W_4 W_3 \mid W_4)$.
\end{enumerate}

\item Winnie-the-Pooh starts at zero of the real line. 
Every minute he moves left by one with probability $0.2$, right by one with probability $0.2$ 
or does not move and eats a honey-pot with probability $0.6$. 


Let $X_t$ be the coordinate of Winnie at time $t$. Rabbit's hole has x-coordinate $6$ and Owl lives at $(-4)$,
let $\tau$ be the first moment when Winnie visits Owl or Rabbit, $\tau = \min\{t \mid X_t = 6 \text{ or } X_t = -4\}$.

\begin{enumerate}
    \item Is $X_t$ a martingale?
    \item Find a constant $a$ such that $Y_t = X_t^2 - at$ is a martingale. 
    \item Find $\P(X_{\tau} = 6)$ and $\E(\tau)$.
\end{enumerate}

Hint: you don't need to check technical conditions of Doob's optional stopping theorem in point (c). 

\item Consider the Vasicek interest rate model 
\[
dR_t = (0.05 - R_t) dt + dW_t, \quad R_0 = 0.07.    
\]
\begin{enumerate}
    \item Write the same models in full form with integrals. 
    \item Let $b_t = \E(R_t)$. Find $b_t$ and sketch it. 
\end{enumerate}

Hint: to find $b_t$ take expected value of both sides of the full form and solve the ordinary differential equation. 

    \item Consider two-period binomial model with initial share price $S_0 = 600$, 
    Up and down multipliers are $u=1.2$, $d=0.8$, risk-free interest rate is $r = 0.05$ per period. 
    
    Consider an option that pays you $X_2 = 100$ at $T=2$ if $S_2 > S_1$ and nothing otherwise. 
    
    \begin{enumerate}
        \item Find the risk neutral probabilities. 
        \item Find the current price $X_0$ of the asset. 
        \item How much shares should I have at $t=1$ in the «down» state of the world to replicate the option?
    \end{enumerate}
    
    
    \item  Consider Black and Scholes model in continuous time with risk-free interest rate $r$, 
    volatility $\sigma$, initial share price $S_0$ and exponential growth rate of expected share price $\mu$. 

    \begin{enumerate}
        \item Let's denote by $\P$ the real probability and by $\P^*$ the risk-neutral probability. 
        Find $\P(S_2 > S_1)$ and $\P^*(S_2 > S_1)$.
        \item An option pays you the sum $X_2 = \sqrt{S_1}$ at $T=2$. 
        Find the non-arbitrage price $X_0$ of the option. 
    \end{enumerate}

    Hint: the answers may contain the standard normal cumulative distribution function $F()$.

    

    
\end{enumerate}


\end{document}
